Discussion Draft

### 2020-04-14 doc/Discussion.rmd
more discussion writing. weird hyper link bug

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>
Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume. This prediction model might fit broader pattern such as in *Arabidopsis* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].

It also could be that the -- greater error in regulation of crossovers may have driven the longer female axis -- to esure / increase the probabilty of succesful maturation of DSB into crossover / chiasmata (cite morelli cohen)

0. There is weak evidence that female chromatin organization having a longer axis is a conserved trait. (only mice and humans, but there is a notable lack of data)

1. We note that A consequence of sexual dmorphism is the chromatin axis (chromatin organization) is that interference strength^norm^ / (general patterns of sex specific genetic maps) will be sexually dimorphic.

2. The density of tethering of loops to the axis is highly conserved (Zand Kleck) - 
variablibilty of the axis length of a given chromosome stems from varing the loop sizes (Z n K)).

3 When interference strength is translated into base pairs (Mb), the chromatin organization with larger loops will have stronger interference (Mb) since more base pairs are sequestered in the DNA loops per unit of SC/axis.

5. while this prediction / the general pattern -- relationhsip of chromatin organization and interference strength is complicated -- the biology of chromatin (3D chromatin organization, heterogeneity of loop size within (cite) and among chromosomes (cite) and (limitations in most / for identifying indiviual chromosomes across single cells)

(this prediction was outlined in Petkov, -- Lynn and Gruhn)

It holds for several model organzism (mice, human, At) and sexual dimorphism in linkage maps (has been one of the oldest measurements of heterochiasmy --  This is an important aspect of how sexual dimorphism in recombination linkage maps has been charaecterized but a lack of cytological evidence for both sexes --

--(so get the cytological data to test if this prediction holds)

<We also hypothesize that the sex differences in chromatin structure are closely connected to sex differences  in the interference strength and our results mirror those in [@petkov2007] and follow the prediction that shorter-loop will have stronger interference.  According to the tethered loop model of DSB formation (cite schmenti?), the 'real estate' for COs are the chromatin / DNA loops, the density of whch is conserved in the axis and loop formation (later the SC str), so the number of loops is equvilent to the length of the axis. **The chromosomes are converted to a 2D space (the axis length of a given chromosome is the limited factor for total CO number)**.  Only one of the models COM aligns with our observed results for sexual dimorphism for axis length. **COM** predicts that axis length will be larger in oocytes due to the larger volume.>

@@

```{r evo.interference, include=FALSE}
#review results,
#why unexpected, unexpected pattern is great for identifying most parimonious model / explanation
#(we're gonna spend alot of space on this section)

# more convincing that the pattern is real

#lil background, for interferences 
#assuptions for predictions (most modifer models, )

#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)

#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)

@@

# consequence of rapid evolution of normalized interference strength,
# in the 3 strains -- larger linkage blocks?
# but what if the rec pattern is repeated?

>>>>

<The results of sexual dimorphism in chromatin structure is supported by the literature, especially for mammals [@lynn2002, @gruhn2013], But there's generally a lack of data for other systems see [@cahoonLibuda2019]).>

Only the COM model predicts sexual dimorphism in chromatin structure will be longer in oocytes due to greater cell volume. This prediction model might fit broader pattern such as in *Arabidopsis thalania* where pollen is the larger cell and has longer axis length [@cahoonLibuda2019].

A consequence of sexual dimorphism chromatin organization is that interference strength quantified in base pairs will be sexually dimorphic. There are two aspects that define this relationship. First the density of tethering of loops to the axis is highly conserved (Zick Kleck), meaning variability of the axis length of a given chromosome comes from variation in the loop sizes (Z n K)).  Second, the signal of crossover interference is mediated through the chromosome axis or SC (kleck Dernberg). Thus the chromatin organization with larger loops (and shorter axis) will have stronger crossover interference when quantified in base pairs, since more base pairs are sequestered in the chromatin loops per unit of axis or SC. Empirical data confirms that interference strength measured in micrometer along the axis is conserved between sexes while interference strength quantified from genetic map is sexually dimorphic, comes from mice [@petkov2007], humans (Gruhn, Lynn). However, due to lack of relevant cytology data in both sexes it is not clear if the many examples of sexual dimorphism in linkage maps is due to chromatin organization.

<a partial pattern seen in Arabidopsis thalinia (Drouaud et al 2007) however there is no SC based measure of interference and low number of double crossovers in the female specific linkage map. It also could be that the greater error in regulation of crossovers may have driven the longer female axis to ensure increase the probability of successful maturation of DSB into crossover chiasmata>

@@

```{r evo.interference, include=FALSE, echo=FALSE}

#models that predict this pattern (CONTROL FOR CO NUMBER -- PRIORITIZE THE WEIGHT /PROB of CO ASSURANCE)
#the pattern is driven in part by short IFD (when short IFD removed in reduced data set -- the low strain mean goes up, but they are still shrter)

@@

# consequence of rapid evolution of normalized interference strength,
# in the 3 strains -- larger linkage blocks?
# but what if the rec pattern is repeated?
```
